---
title: "Decision Making with Sentiment Analysis in R"
subtitle: "From Social Media Noise to Managerial Action"
author: "Mustafa Asım Ruhi 401190044<br>ruhi.mustafa@stud.hs-fresenius.de</span>"
institute: "Fresenius University of Applied Sciences<br>Course: Data Analysis for Decision-Making<br>Professor: Prof. Dr. Stephan Huber<br>2 June 2026"
date: ""
format:
  revealjs:
    theme: simple
    width: 1280
    height: 720
    margin: 0.10
    center: true
    slide-number: true
    embed-resources: true
    toc: false
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup
#| include: false

suppressPackageStartupMessages({
  library(tidyverse)
  library(tidytext)
  library(janitor)
  library(lubridate)
  library(scales)
  library(knitr)
})

url <- "https://raw.githubusercontent.com/ruchitgandhi/Twitter-Airline-Sentiment-Analysis/master/Tweets.csv"

tweets <- read_csv(url, show_col_types = FALSE) |>
  clean_names() |>
  select(
    airline_sentiment,
    negativereason,
    airline,
    retweet_count,
    text,
    tweet_created
  ) |>
  mutate(
    tweet_created = parse_date_time(tweet_created, orders = c("ymd HMS z", "ymd HMS")),
    airline_sentiment = factor(
      airline_sentiment,
      levels = c("negative", "neutral", "positive")
    )
  )

sentiment_by_airline <- tweets |>
  count(airline, airline_sentiment) |>
  group_by(airline) |>
  mutate(share = n / sum(n)) |>
  ungroup()

negative_reasons <- tweets |>
  filter(airline_sentiment == "negative", !is.na(negativereason)) |>
  count(negativereason, sort = TRUE)

tweets_words <- tweets |>
  mutate(tweet_id = row_number()) |>
  select(tweet_id, airline, airline_sentiment, text) |>
  unnest_tokens(word, text) |>
  anti_join(stop_words, by = "word") |>
  filter(
    str_detect(word, "[a-z]"),
    !str_detect(word, "^http"),
    !str_detect(word, "^t.co")
  )

word_sentiments <- tweets_words |>
  inner_join(get_sentiments("bing"), by = "word")

top_sentiment_words <- word_sentiments |>
  count(sentiment, word, sort = TRUE) |>
  group_by(sentiment) |>
  slice_max(n, n = 8) |>
  ungroup()

lexicon_airline_sentiment <- word_sentiments |>
  count(airline, sentiment) |>
  pivot_wider(
    names_from = sentiment,
    values_from = n,
    values_fill = 0
  ) |>
  mutate(
    lexicon_score = positive - negative,
    total_sentiment_words = positive + negative,
    lexicon_score_per_100_words = 100 * lexicon_score / total_sentiment_words
  ) |>
  arrange(lexicon_score_per_100_words)

decision_actions <- negative_reasons |>
  mutate(
    managerial_action = case_when(
      negativereason == "Customer Service Issue" ~ "Improve support capacity and service recovery.",
      negativereason == "Late Flight" ~ "Improve delay communication and punctuality.",
      negativereason == "Cancelled Flight" ~ "Strengthen rebooking and compensation.",
      negativereason == "Lost Luggage" ~ "Improve baggage tracking.",
      negativereason == "Bad Flight" ~ "Improve in-flight experience.",
      TRUE ~ "Investigate this category further."
    )
  )
```

## Research Question

<div class="hero-text">
How can customer tweets be transformed into decision-relevant insights for airline management?
</div>

## What is Sentiment Analysis?

<div class="main-text">
Sentiment analysis is a text mining method used to identify opinions, emotions, and attitudes in written language.
</div>

<div class="three-boxes">
<div class="box positive">Positive</div>
<div class="box neutral">Neutral</div>
<div class="box negative">Negative</div>
</div>

## Why It Matters for Decision-Making

<div class="three-boxes">
<div class="box">Companies receive large amounts of customer feedback.</div>
<div class="box">Managers need fast and structured insights.</div>
<div class="box">Sentiment analysis helps prioritize business actions.</div>
</div>

## Different Sentiment Analysis Methods

| Method | Basic idea | Why not used / selected? |
|---|---|---|
| Lexicon-based | Uses predefined sentiment dictionaries | Selected because it is transparent and easy to reproduce in RStudio |
| Machine learning | Trains a model on labeled text data | Requires training/testing setup |
| Transformer-based | Uses context-aware models such as BERT | More complex for a short live demo |
| Hybrid | Combines dictionaries and machine learning | More advanced setup |


In this project, I use a lexicon-based approach because it is transparent, reproducible, and suitable for a short RStudio demonstration.

## From Text to Managerial Action

<div class="flow-grid">
<div class="flow-box"><span class="step-number">1</span><span class="step-label">Collect text</span></div>
<div class="flow-box"><span class="step-number">2</span><span class="step-label">Clean data</span></div>
<div class="flow-box"><span class="step-number">3</span><span class="step-label">Measure sentiment</span></div>
<div class="flow-box"><span class="step-number">4</span><span class="step-label">Find problems</span></div>
<div class="flow-box"><span class="step-number">5</span><span class="step-label">Recommend action</span></div>
</div>

## Case Study: Airline Tweets

<div class="main-text">
The example uses <b>14,640 airline-related tweets</b> classified as negative, neutral, or positive.
</div>

```{r}
tweets |>
  count(airline_sentiment) |>
  mutate(share = percent(n / sum(n))) |>
  kable(col.names = c("Sentiment", "Tweets", "Share"))
```

## Sentiment by Airline

```{r}
#| fig-width: 8.5
#| fig-height: 4.3

ggplot(sentiment_by_airline, aes(x = airline, y = share, fill = airline_sentiment)) +
  geom_col(position = "fill") +
  scale_y_continuous(labels = percent) +
  labs(x = "Airline", y = "Share of tweets", fill = "Sentiment") +
  theme_minimal(base_size = 15) +
  theme(legend.position = "bottom")
```

## Main Negative Reasons

```{r}
#| fig-width: 8.5
#| fig-height: 4.3

negative_reasons |>
  slice_max(n, n = 8) |>
  ggplot(aes(x = reorder(negativereason, n), y = n)) +
  geom_col() +
  coord_flip() +
  labs(x = "Negative reason", y = "Number of tweets") +
  theme_minimal(base_size = 15)
```

## R Workflow: Tokenization

```{r}
#| echo: true
#| eval: false

tweets_words <- tweets |>
  mutate(tweet_id = row_number()) |>
  select(tweet_id, airline, airline_sentiment, text) |>
  unnest_tokens(word, text) |>
  anti_join(stop_words, by = "word")
```

## Lexicon-Based Sentiment Analysis

```{r}
#| echo: true
#| eval: false

word_sentiments <- tweets_words |>
  inner_join(get_sentiments("bing"), by = "word")
```

## Frequent Sentiment Words

```{r}
#| fig-width: 8.5
#| fig-height: 4.3

top_sentiment_words |>
  mutate(word = reorder(word, n)) |>
  ggplot(aes(x = word, y = n, fill = sentiment)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ sentiment, scales = "free") +
  coord_flip() +
  labs(x = "Word", y = "Frequency") +
  theme_minimal(base_size = 15)
```

## Sentiment Score by Airline

```{r}
#| fig-width: 8.5
#| fig-height: 4.3

ggplot(
  lexicon_airline_sentiment,
  aes(
    x = reorder(airline, lexicon_score_per_100_words),
    y = lexicon_score_per_100_words
  )
) +
  geom_col() +
  coord_flip() +
  labs(x = "Airline", y = "Sentiment score per 100 words") +
  theme_minimal(base_size = 15)
```

## From Analysis to Decisions

```{r}
decision_actions |>
  select(negativereason, n, managerial_action) |>
  head(5) |>
  kable(col.names = c("Problem", "Tweets", "Managerial action"))
```

## Managerial Interpretation

<div class="hero-text">
The real value is not only knowing whether customers are happy or unhappy.
</div>

<div class="main-text">
The real value is knowing which operational problem should be solved first.
</div>

## Strengths and Limitations

<div class="two-boxes">
<div class="box">
<b>Strengths</b><br><br>
Fast analysis<br>
Reproducible workflow<br>
Useful for monitoring
</div>

<div class="box">
<b>Limitations</b><br><br>
Sarcasm is difficult<br>
Context can be missed<br>
Human interpretation is needed
</div>
</div>

## In-Class Exercise

<div class="main-text">
Students analyze three short customer comments in R and recommend one managerial action.
</div>

```{r}
#| echo: true
#| eval: false

customer_comments <- tibble(
  text = c(
    "The flight was delayed and the service was terrible",
    "The staff was helpful and the flight was comfortable",
    "My luggage was lost and nobody helped me"
  )
)

customer_comments |>
  unnest_tokens(word, text) |>
  anti_join(stop_words, by = "word") |>
  inner_join(get_sentiments("bing"), by = "word") |>
  count(sentiment)
```

## Conclusion

<div class="hero-text">
Sentiment analysis turns unstructured customer feedback into practical decision support.
</div>

<div class="main-text">
For managers, the key question is not only: <b>“What is the sentiment?”</b><br>
It is: <b>“What decision should this sentiment trigger?”</b>
</div>

## Thank You

<div class="main-text">
<b>Mustafa Asım Ruhi</b><br>
Student number: 401190044<br>
Data Analysis for Decision-Making<br>
2 June 2026
</div>

<style>
.reveal {
  font-size: 33px;
}

.reveal .slides {
  text-align: center;
}

.reveal .slides section {
  box-sizing: border-box;
  padding: 45px 85px;
}

.reveal h1 {
  font-size: 1.65em;
}

.reveal h2 {
  font-size: 1.35em;
  margin-bottom: 0.55em;
}

.small-info {
  font-size: 0.55em;
  line-height: 1.25;
}

.main-text {
  max-width: 900px;
  margin-left: auto;
  margin-right: auto;
  font-size: 0.95em;
  line-height: 1.35;
}

.hero-text {
  max-width: 930px;
  margin-left: auto;
  margin-right: auto;
  font-size: 1.15em;
  line-height: 1.28;
  font-weight: 700;
}

.three-boxes,
.two-boxes,
.flow-grid {
  display: grid;
  gap: 24px;
  margin-left: auto;
  margin-right: auto;
  align-items: stretch;
  justify-content: center;
}

.three-boxes {
  grid-template-columns: repeat(3, 1fr);
  max-width: 980px;
  width: 90%;
  margin-top: 35px;
}

.two-boxes {
  grid-template-columns: repeat(2, 1fr);
  max-width: 900px;
  width: 88%;
}

.flow-grid {
  grid-template-columns: repeat(5, 1fr);
  max-width: 1050px;
  width: 94%;
}

.box,
.flow-box {
  border: 2px solid #d7e7dc;
  background: #f3fbf5;
  border-radius: 18px;
  padding: 24px 18px;
  min-height: 110px;
  display: flex;
  align-items: center;
  justify-content: center;
  line-height: 1.2;
  box-sizing: border-box;
}

.flow-box {
  font-size: 0.68em;
  min-height: 125px;
  padding: 20px 12px;
}

.positive,
.neutral,
.negative {
  font-weight: 700;
}

.reveal table {
  margin-left: auto;
  margin-right: auto;
  font-size: 0.58em;
  max-width: 95%;
}

.reveal pre {
  width: 88%;
  margin-left: auto;
  margin-right: auto;
}

.reveal pre code {
  font-size: 0.56em;
  line-height: 1.15;
  max-height: 420px;
}

.reveal img,
.reveal svg {
  max-width: 90%;
  max-height: 485px;
  object-fit: contain;
}

.reveal .slide-number {
  font-size: 0.45em;
}
</style>

<style>
.flow-box { flex-direction: column !important; gap: 12px !important; text-align: center !important; }
.step-number { display: block; font-size: 1.35em; font-weight: 800; line-height: 1; }
.step-label { display: block; font-size: 0.92em; line-height: 1.15; }
</style>

<style id='flow-align-safe-fix'>
.flow-box {
  display: flex !important;
  flex-direction: column !important;
  justify-content: center !important;
  align-items: center !important;
  text-align: center !important;
  gap: 8px !important;
}
.flow-box b {
  display: block !important;
  font-size: 1.25em !important;
  line-height: 1 !important;
  margin: 0 !important;
}
</style>

